Appendix

### 2021-05-07 README.Rmd
update readme.md

---
title: A Road Segment Prioritization Approach for Cycling Infrastructure
#date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # # bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options: 
  markdown: 
    wrap: 72
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: elsevier-harvard.csl
---

@@

<!-- # Ideas and discussion -->

<!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->

<!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->

<!-- Can go into road segment section. -->

<!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->

<!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->

<!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->

<!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->

<!-- Keywords -->

@@

Our work builds on these *community finding* approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose an approach that
accounts for motivators and deterrents to cycling. We propose three
sub-methods that address some of the limitations of previous
studies. These limitations include (a) an inherent bias of basing network
design solely on existing cycling demand, (b) the proposal of routes that may
not correspond to studies on cyclist preference and government policies,
and (c) an insufficient consideration of the ethical principles underlying
the analysis. The current work addresses these limitations through the following organisation. Section \@ref(calculating-potential-cycling-demand)
focuses on calculating potential cycling demand. Section \@ref(routing)
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\@ref(road-segment-prioritisation) outlines a method for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'.

@@

Many of the cycling network studies mentioned above use demand data for cycling as a starting point. 
Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
@duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
entire study area. 
@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
the potential demand distribution to the distribution obtained from GPS data.

@@

# Road Segment Prioritisation

@@

A range of algorithms could be used for prioritisation. Because policy
priorities vary, we present two algorithms. Both utilize existing
infrastructure from the beginning and allow us to compare a solution
that focuses on utilitarianism to one that focuses on egalitarianism. In
both algorithms, links are selected iteratively and the iteration at
which each link is added to the solution is recorded. Investments in
cycling infrastructure can be limited by budget constraints, so it can
be useful to see where best to allocate a defined length of segregated
infrastructure. In order to incorporate egalitarian principles in our
approach, we use community detection to partition the study area and
evaluate the distribution of investment over the different subdivisions.

@@

Given that the "most essential activity entailed in the design of
cycle-friendly infrastructure is developing a cycle network"
[@parkin2018designing], we believe that the approach has great
potential to inform investment in cities such as Manchester where there
is political will to invest in cycling long-term. A benefit of the
approach is that it has relatively modest data requirements: only the
road network (from OSM), its topography (from satellite imagery), and
OD data (from the national census in this case) are
needed, opening up the possibility of deploying the approach in other
cities^[The results are easily reproducible for all UK cities, and can
also be reproduced for cities elsewhere given the availability of
commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham. 
For instructions on reproducing the results shown in this
paper, see the **README** file in the R folder of the .zip file that
accompanies this paper (to be released as open source software when the
paper is published).].

@@

The approach can identify not only where there is high
potential for cycling but also trade-offs between
stress levels (related to motor traffic) and directness.
The results can therefore be used as a basis for recommendations on
road space reallocation *and* new infrastructure to unlock potential cycling demand.
The approach encourages consideration of a wider range of preferences
and needs than previous approaches that focus only on absolute potential.
Moreover, the inclusion of egalitarian principles in scenarios of change
encourages investment in cycling infrastructure to
increase the connectivity of existing cycling infrastructure
*and* investment that addresses geographical and social inequalities.
This ability to address inequalities in network prioritisation is particularly
important given research showing substantial *inequalities*
around transport provision in general and cycling uptake and investment
in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].

>>>>

---
title: A Road Segment Prioritization Approach for Cycling Infrastructure
#date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
 # bookdown::github_document2:
   bookdown::pdf_document2:
 # bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options:
  markdown:
    wrap: 72
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: elsevier-harvard.csl
always_allow_html: yes
---

@@

@@

Our work builds on these *community finding* approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose an approach that
accounts for motivators and deterrents to cycling. We propose three
sub-methods that address some of the limitations of previous
studies. These limitations include (a) an inherent bias of basing network
design solely on existing cycling demand, (b) the proposal of routes that may
not correspond to studies on cyclist preference and government policies,
and (c) an insufficient consideration of the ethical principles underlying
the analysis. The current work addresses these limitations through the following organisation. Section \@ref(calculating-potential-cycling-demand)
focuses on calculating potential cycling demand. Section \@ref(routing)
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\@ref(road-segment-prioritization) outlines a method for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'.

@@

Many of the cycling network studies mentioned above use demand data for cycling as a starting point. 
Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
@duthie2014optimization note that relying only on existing cycling activity to prioritize cycling
infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
entire study area. 
@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
the potential demand distribution to the distribution obtained from GPS data.

@@

# Road Segment Prioritization

@@

A range of algorithms could be used for prioritization. Because policy
priorities vary, we present two algorithms. Both utilize existing
infrastructure from the beginning and allow us to compare a solution
that focuses on utilitarianism to one that focuses on egalitarianism. In
both algorithms, links are selected iteratively and the iteration at
which each link is added to the solution is recorded. Investments in
cycling infrastructure can be limited by budget constraints, so it can
be useful to see where best to allocate a defined length of segregated
infrastructure. In order to incorporate egalitarian principles in our
approach, we use community detection to partition the study area and
evaluate the distribution of investment over the different subdivisions.

@@

Given that the "most essential activity entailed in the design of
cycle-friendly infrastructure is developing a cycle network"
[@parkin2018designing], we believe that the approach has great
potential to inform investment in cities such as Manchester where there
is political will to invest in cycling long-term. A benefit of the
approach is that it has relatively modest data requirements: only the
road network (from OSM), its topography (from satellite imagery), and
OD data (from the national census in this case) are
needed, opening up the possibility of deploying the approach in other
cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]

<!-- [The results are easily reproducible for all UK cities, and can -->
<!-- also be reproduced for cities elsewhere given the availability of -->
<!-- commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham.  -->
<!-- For instructions on reproducing the results shown in this -->
<!-- paper, see the **README** file in the R folder of the .zip file that -->
<!-- accompanies this paper (to be released as open source software when the -->
<!-- paper is published).]. -->

@@

The approach can identify not only where there is high
potential for cycling but also trade-offs between
stress levels (related to motor traffic) and directness.
The results can therefore be used as a basis for recommendations on
road space reallocation *and* new infrastructure to unlock potential cycling demand.
The approach encourages consideration of a wider range of preferences
and needs than previous approaches that focus only on absolute potential.
Moreover, the inclusion of egalitarian principles in scenarios of change
encourages investment in cycling infrastructure to
increase the connectivity of existing cycling infrastructure
*and* investment that addresses geographical and social inequalities.
This ability to address inequalities in network prioritization is particularly
important given research showing substantial *inequalities*
around transport provision in general and cycling uptake and investment
in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].

### 2021-05-07 appendix.Rmd
update readme.md

---
title: "Appendix"
author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
date: "13/03/2021"
geometry: margin=1.5cm
output:
  pdf_document:
    extra_dependencies: ["subcaption", "float"]
---

>>>>

---
title: "Appendix"
#author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
#date: "13/03/2021"
geometry: margin=1.5cm
output:
  pdf_document:
    extra_dependencies: ["subcaption", "float"]
---
